---
title: "M2_Lab"
author: "Emily Shepherd"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I created two different histograms. For the first one, I basically created the same histogram in the data, but with a different variable. I used the built-in histogram feature.

For the second histogram, I used the data from the lab. I changed the binwidth to 2. I also changed the colors of bars. The outline is now a navy blue and the fill is gray. Additionally, I decreased the transparency. Lastly, I loaded the ggthemes library and used the theme_economist to give it a more professional polish.

```{r, echo = TRUE, warning=FALSE}
setwd("/Users/burrisfaculty/Desktop/DSCode/DSCI605")
library(ggplot2)
library(ggthemes)
soc = read.csv(file = 'soil organic carbon.csv')
x5cm <- soc[,2]
x510cm <- soc[,3]
hist(x510cm)

ggplot(data = NULL, aes(x5cm)) + 
  geom_histogram(breaks = seq(0, max(x5cm), by = 2),
                 col = 'navy',
                 fill = 'gray',
                 alpha = 0.6) + 
  labs(title = 'Histogram for SOC', x = 'SOC(%)', y = 'Count') +
  xlim(c(0, max(x5cm))) + 
  theme_economist()
```
